\section{Summary}
\subsection{Introduction}
The Stahel-Donoho estimator is a well-known affine equivariant robust estimator
of multivariate location and scatter.
The estimator gains its robustness by using a weighting function, giving a
small weight to observations which have a large ``outlyingness.''
However the SD estimator has difficulties in identifying the correct
outlyingness when a majority of the observations are contaminated in at least
one of their components.
This is a problem because contamination of components is common in data with
a high number of dimensions.
By huberizing the data before calculating the outlyingness, which yields the
huberized outlyingness the robustness of the estimator towards such
contaminated data is increased.

\subsection{Contamination Model}
The standard contamination model assumes that the majority of the observations
are regular, following some underlying model distribution, while the
remaining minority, the outliers, can be anything.
For high-dimensional data this model is unrealistic, it is far more likely that
most observations have been contaminated in some of their components.
A more flexible model is needed to study the influence of contamination in
many dimensions.
Let $X$, $Y$, and $Z$ be $p$-dimensional random vectors, where $Y$ follows
some regular distribution and $Z$ has an arbitrary distribution that generates
the outliers.
Then the observed random variable $X$ follows the model
    \[ X = (\mathbf{I} - \mathbf{B})Y + \mathbf{B}Z, \]
where $\mathbf{B} = diag(\mathit{B}_1, \mathit{B}_2, \dots, \mathit{B}_p)$
is a diagonal matrix and $\mathit{B}_1, \mathit{B}_2, \dots, \mathit{B}_p$
are Bernoulli random variables with $P(\mathit{B}_i = 1) = \epsilon_i$.

The standard contamination model corresponds to the assumption
$P(\mathit{B}_1 = \mathit{B}_2 = \dots = \mathit{B}_p) = 1$.
Such contaminated observations are called \emph{structural outliers}.
These are not very interesting because the regular SD estimator copes very
well with them and improvements are not necessary/expected for this type of
outlier.

The case considered further is that of \emph{componentwise outliers} which
corresponds to the assumption that
$\mathit{B}_1, \mathit{B}_2, \dots, \mathit{B}_p$ are independent.
If $P(\mathit{B}_i = 1) = \epsilon$ for all components, then each variable contains
on average a fraction $(1-\epsilon)$ of clean measurements, but the probability
that an observation is completely uncontaminated is only $(1-\epsilon)^p$.
Even for moderate $\epsilon$ this probability quickly exceeds $50\%$ for high
dimensions $p$, so there is no majority of outlier-free observations.
The independent model does not support affine transformations because of
\emph{outlier propagation}.
As a consequence, affine equivariant estimators are not very robust under this
model.

\subsection{Stahel-Donoho estimator}
Let $\mathbf{X}$ be an $n \times p$ data matrix that contains $n$ observations
$x_1, \dots, x_n$ in $\mathbb{R}^p$.
Let $\mu$ and $\sigma$ be shift and scale equivariant univariate location and
scale statistics.
Then, for any $y \in \mathbb{R}^p$, the Stahel-Donoho \emph{outlyingness} is
defined as
    \[ r(y, \mathbf{X}) = \sup_{a \in S_p} \frac{|y' a - \mu(\mathbf{X}a)|}
        {\sigma(\mathbf{X}a)}, \]
with $S_p = \{ a \in \mathbb{R}^p : \|a\| = 1 \}$.
From now on, we will denote $r(x_i, \mathbf{X})$ by $r_i$.
The Stahel-Donoho estimator of location and scatter $(T_{SD}, \mathbf{S}_{SD})$
is defined as
    \[ T_{SD} = \frac{\sum\limits_{i=1}^n w_i x_i}{\sum\limits_{i=1}^n w_i}, \]
and
    \[ \mathbf{S}_{SD} = \frac{\sum\limits_{i=1}^n w_i (x_i - T_{SD})(x_i - T_{SD})'}
        {\sum\limits_{i=1}^n w_i}, \]
where $w_i = w(r_i)$ is a weight function so that observations with large
outlyingness get small weights.
To attain the maximum breakdown point $\mu$ is taken to be the median and
$\sigma$ is chosen to be the modified median absolute deviation.

From these definitions it is clear that the SD estimators are a weighted mean
and covariance matrix.
The weight depends on the outlyingness, so the robustness of the estimator
primarily depends on the calculation of the outlyingness.

In figures \ref{fig:1} through \ref{fig:3} the problems with this calculation
of the outlyingness are highlighted.
The outliers that are outlying in both dimensions are masked because the
majority of the observations are outlying in at least one  dimension and the
regular observations are swamped because of this.
The SD estimator can clearly not cope with this kind of contamination.

\subsection{Stahel-Donoho estimator with huberized outlyingness (HSD)}
To improve the robustness of the SD estimator the calculation of the
outlyingness is adapted.
To avoid the masking and swamping effects of the outliers the data is first
huberized before the median and MAD are calculated.
Then the outlyingness is calculated using these location and scale estimates.
    \[ r_H(y,\mathbf{X}) = \sup_{a \in S_p} \frac{|y'a - \mu(\mathbf{X}_H a)|}
        {\sigma{\mathbf{X}_H a}} \]
Becaus of the huberization the outliers have far less influence on the
estimators for scale and location used in the calculation of the outlyingness
and consequently their outlyingess is far greater.

The resulting HSD estimator loses the affine equivariance property because the
huberization is a componentwise transformation.
However, the estimator is far more robust in situations with many contaminated
observations, which is realistic in real high-dimensional datasets.
